---
title: "Practice Exercises in R"
author: "Connor Birkhold"
date: "May 23, 2019"
output: html_document
---

```{r include=FALSE}
setwd("C:/Users/Cbirkho/Documents/Practice_Exercises")
```


#### Load Packages
```{r warning=FALSE,message=FALSE}
library(tidyverse)
library(readxl)
```
#### Load in the data
```{r warning=FALSE}
d_customer <- read_excel('./data/tables.xlsx',
                         sheet = 'd_customer') %>% 
  mutate(signup_date = as.Date(as.numeric(signup_date), origin = '1899-12-30'))

line_sales <- read_excel('./data/tables.xlsx',
                         sheet = 'line_sales') %>% 
  rename(purchase_channel = `purchase channel`)

product_table <- read_excel('./data/tables.xlsx',
                            sheet = 'product_table')

```
#### 1. Write code to provide a list of clients that purchased the brand 'gap' and have an email address
```{r}

newTable <- left_join(line_sales, product_table, by = 'product_key')
newTable <- left_join(newTable, d_customer, by = 'customer_id')

newTable <- newTable %>% 
  filter(brand == 'gap' & !is.na(email_address)) %>% 
  select(customer_id, email_address)

print(newTable)

```
#### 2. Write code to provide three lists.  

* a.One list of distinct clients that purchased from only the online channel and not the store channel.
```{r}

filterTable <- line_sales %>% 
  filter(purchase_channel == 'online') %>% 
  select(customer_id)

newTable <- line_sales %>% 
  filter(customer_id %in% filterTable$customer_id) %>% 
  group_by(customer_id) %>% 
  summarise(cnt_purchase_channel = length(purchase_channel)) %>% 
  filter(cnt_purchase_channel == 1) %>% 
  select(customer_id)

print(newTable)

```
* b.One list of distinct clients that purchase from only the store channel and not online
```{r}
filterTable <- line_sales %>% 
  filter(purchase_channel == 'store') %>% 
  select(customer_id)

newTable <- line_sales %>% 
  filter(customer_id %in% filterTable$customer_id) %>% 
  group_by(customer_id) %>% 
  summarise(cnt_purchase_channel = length(purchase_channel)) %>% 
  filter(cnt_purchase_channel == 1) %>% 
  select(customer_id)

print(newTable)

```
* c.One list of distinct clients that purchased from both the online and store channel.
```{r}
onlineTable <- line_sales %>% 
  filter(purchase_channel == 'online') %>% 
  select(customer_id)

storeTable <- line_sales %>% 
  filter(purchase_channel == 'store') %>% 
  select(customer_id)

newTable <- onlineTable %>% 
  filter(customer_id %in% storeTable$customer_id)
  

print(newTable)

```

#### 3. Write code to provide the count of distinct clients, total orders, total sales, and average purchase by purchase channel
```{r warning=FALSE, message=FALSE}
library(data.table) # going to use data.table package here
```

```{r}


setDT(line_sales) # converting line_sales to data.table structure

newTable <- line_sales[,
                       .(distinct_clients = uniqueN(customer_id),
                         total_orders = uniqueN(order_number),
                         total_sales = sum(total_amt),
                         avg_purchase = mean(total_amt)),
                       by = purchase_channel]


print(newTable)
```

#### 4.Write code to provide a list of distinct clients that have an email address, signed up for email after 01/01/2010, and purchased tops after 3/10/2010
```{r}
newTable <- left_join(d_customer, line_sales, by = 'customer_id')

newTable <- left_join(newTable, product_table, by = 'product_key')

newTable <- newTable %>% 
  filter(!is.na(email_address)
         & class == 'tops'
         & signup_date > '2010-01-01'
         & order_date > '2010-03-10') %>% 
  select(customer_id)

print(newTable) # no customer_id meet these conditions


```

#### 5.Write code to provide a list of each email address and each email addresses total sales.
```{r}
newTable <- left_join(d_customer, line_sales, by = 'customer_id')

newTable <- newTable %>% 
  filter(!is.na(email_address)) %>% 
  group_by(email_address) %>% 
  summarise(total_sales = sum(total_amt)) %>% 
  filter(!is.na(total_sales))

print(newTable)

```

