---
title: "Practice Exercises SQL"
author: "Connor Birkhold"
date: "May 25, 2019"
output: html_document
---

```{r include=FALSE}
setwd("C:/Users/Cbirkho/Documents/Practice_Exercises")
```


#### Load Packages
```{r warning = FALSE, message = FALSE}
library(sqldf) # package used to write SQL in R
library(tidyverse)
library(readxl)

```

#### Load the data
```{r warning=FALSE}
d_customer <- read_excel('./data/tables.xlsx',
                         sheet = 'd_customer') %>% 
  mutate(signup_date = as.Date(as.numeric(signup_date), origin = '1899-12-30'))

line_sales <- read_excel('./data/tables.xlsx',
                         sheet = 'line_sales') %>% 
  rename(purchase_channel = `purchase channel`)

product_table <- read_excel('./data/tables.xlsx',
                            sheet = 'product_table')

```

#### 1. Write code to provide a list of clients that purchased the brand 'king' and have an email address
```{r}
sqldf("SELECT 
            ls.customer_id, 
            c.email_address
        FROM product_table pt 
        JOIN line_sales ls 
          ON pt.product_key = ls.product_key
        JOIN d_customer c 
          ON ls.customer_id = c.customer_id
        WHERE pt.brand = 'king'
        AND c.email_address IS NOT NULL;")


```

#### 2. Write code to provide three lists.  

* A.One list of distinct clients that purchased from only the online channel and not the store channel.
```{r}
sqldf("SELECT 
        ls1.customer_id 
      FROM line_sales ls1 
      JOIN (SELECT
              ls2.customer_id 
            FROM line_sales ls2 
            WHERE purchase_channel = 'online') op 
        ON ls1.customer_id = op.customer_id
      GROUP BY ls1.customer_id
      HAVING COUNT(ls1.purchase_channel) = 1;")

# Solve 2A using CTE
sqldf("WITH
      online_sale AS (SELECT 
                        customer_id
                      FROM line_sales
                      WHERE purchase_channel = 'online'),
      store_sale AS (SELECT customer_id
                     FROM line_sales
                     WHERE purchase_channel = 'store')
      SELECT 
        os.customer_id
      FROM online_sale os
      LEFT JOIN store_sale ss 
      ON os.customer_id = ss.customer_id
      WHERE ss.customer_id IS NULL;")

```

* B.One list of distinct clients that purchase from only the store channel and not online
```{r}
sqldf("SELECT 
        ls1.customer_id 
      FROM line_sales ls1 
      JOIN (SELECT
              ls2.customer_id 
            FROM line_sales ls2 
            WHERE purchase_channel = 'store') op 
        ON ls1.customer_id = op.customer_id
      GROUP BY ls1.customer_id 
      HAVING COUNT(ls1.purchase_channel) = 1;") 

# Solve 2B using CTE
sqldf("WITH
      online_sale AS (SELECT 
                        customer_id
                      FROM line_sales
                      WHERE purchase_channel = 'online'),
      store_sale AS (SELECT 
                        customer_id
                     FROM line_sales
                     WHERE purchase_channel = 'store')
      SELECT
        ss.customer_id
      FROM store_sale ss
      LEFT JOIN online_sale os 
      ON ss.customer_id = os.customer_id
      WHERE os.customer_id IS NULL;")

```

* C.One list of distinct clients that purchased from both the online and store channel.
```{r}
sqldf("SELECT
        ls1.customer_id 
      FROM line_sales ls1
      JOIN (SELECT 
              DISTINCT customer_id 
            FROM line_sales 
            WHERE purchase_channel <> 'online') v0 
        ON ls1.customer_id = v0.customer_id
      WHERE ls1.purchase_channel = 'online'; ")

# Solve 2C using CTE
sqldf("WITH
      online_sale AS (SELECT
                        customer_id
                      FROM line_sales
                      WHERE purchase_channel = 'online'),
      store_sale AS (SELECT
                        customer_id
                     FROM line_sales
                     WHERE purchase_channel = 'store')
      SELECT
        ss.customer_id
      FROM store_sale ss
      JOIN online_sale os 
      ON ss.customer_id = os.customer_id
      WHERE os.customer_id IS NOT NULL
      AND ss.customer_id IS NOT NULL
      GROUP BY 
        ss.customer_id;")

```

#### 3. Write code to provide the count of distinct clients, total orders, total sales, and average purchase by purchase channel
```{r}

sqldf("SELECT 
          purchase_channel
          ,COUNT(DISTINCT customer_id) AS distinct_clients
          ,COUNT(DISTINCT order_number) AS total_orders
          ,SUM(total_amt) AS total_sales
          ,AVG(total_amt) AS avg_purchase
      FROM line_sales 
      GROUP BY
        purchase_channel;")



```

#### 4.Write code to provide a list of distinct clients that have an email address, signed up for email after 01/01/2010, and purchased tops after 3/10/2010
**Note: sqldf package in R does not like the date filters. This will run in a SQL editor**
```{r error=TRUE, eval=FALSE}


sqldf("SELECT
        c.customer_id
      FROM d_customer c
      JOIN line_sales ls 
        ON c.customer_id = ls.customer_id
      JOIN product_table pt
        ON ls.product_key = pt.product_key
      WHERE c.email_address is not null 
      AND c.signup_date > '2010-01-01'
      AND ls.order_date > '2010-03-10'
      AND pt.class = 'tops';")

```
#### 5.Write code to provide a list of each email address and each email addresses total sales.
```{r}


sqldf("SELECT 
        c.email_address, 
        SUM(ls.total_amt) AS total_sales
      FROM d_customer c 
      JOIN line_sales ls
        ON c.customer_id = ls.customer_id
      WHERE c.email_address IS NOT NULL
      GROUP BY
        c.email_address")
```
