---
title: "Practice Exercises SQL"
author: "Connor Birkhold"
date: "May 25, 2019"
output: html_document
---

```{r include=FALSE}
setwd("C:/Users/Cbirkho/Documents/Practice_Exercises")
```


#### Load Packages
```{r warning = FALSE, message = FALSE}
library(sqldf) # package used to write SQL in R
library(tidyverse)
library(readxl)

```

#### Load the data
```{r warning=FALSE}
d_customer <- read_excel('./data/tables.xlsx',
                         sheet = 'd_customer') %>% 
  mutate(signup_date = as.Date(as.numeric(signup_date), origin = '1899-12-30'))

line_sales <- read_excel('./data/tables.xlsx',
                         sheet = 'line_sales') %>% 
  rename(purchase_channel = `purchase channel`)

product_table <- read_excel('./data/tables.xlsx',
                            sheet = 'product_table')

```

#### 1. Write code to provide a list of clients that purchased the brand 'gap' and have an email address
```{r}
sqldf("select 
            ls.customer_id, 
            c.email_address
        from product_table pt 
        join line_sales ls 
          on pt.product_key = ls.product_key
        join d_customer c 
          on ls.customer_id = c.customer_id
        where pt.brand = 'gap'
        and c.email_address is not null")


```

#### 2. Write code to provide three lists.  

* A.One list of distinct clients that purchased from only the online channel and not the store channel.
```{r}
sqldf("select 
        ls1.customer_id 
      from line_sales ls1 
      join (select ls2.customer_id 
            from line_sales ls2 
            where purchase_channel = 'online') op 
        on ls1.customer_id = op.customer_id
      group by ls1.customer_id
      having count(ls1.purchase_channel) = 1")

# Solve 2A using CTE
sqldf("with
      online_sale as (select customer_id
                      from line_sales
                      where purchase_channel = 'online'),
      store_sale as (select customer_id
                     from line_sales
                     where purchase_channel = 'store')
      select os.customer_id
      from online_sale os
      left join store_sale ss 
      on os.customer_id = ss.customer_id
      where ss.customer_id is null")

```

* B.One list of distinct clients that purchase from only the store channel and not online
```{r}
sqldf("select 
        ls1.customer_id 
      from line_sales ls1 
      join (select ls2.customer_id 
            from line_sales ls2 
            where purchase_channel = 'store') op 
        on ls1.customer_id = op.customer_id
      group by ls1.customer_id 
      having count(ls1.purchase_channel) = 1") 

# Solve 2B using CTE
sqldf("with
      online_sale as (select customer_id
                      from line_sales
                      where purchase_channel = 'online'),
      store_sale as (select customer_id
                     from line_sales
                     where purchase_channel = 'store')
      select ss.customer_id
      from store_sale ss
      left join online_sale os 
      on ss.customer_id = os.customer_id
      where os.customer_id is null")

```

* C.One list of distinct clients that purchased from both the online and store channel.
```{r}
sqldf("select 
        ls1.customer_id 
      from line_sales ls1
      join (select distinct customer_id 
            from line_sales 
            where purchase_channel <> 'online') v0 
        on ls1.customer_id = v0.customer_id
      where ls1.purchase_channel = 'online' ")

# Solve 2C using CTE
sqldf("with
      online_sale as (select customer_id
                      from line_sales
                      where purchase_channel = 'online'),
      store_sale as (select customer_id
                     from line_sales
                     where purchase_channel = 'store')
      select ss.customer_id
      from store_sale ss
      join online_sale os 
      on ss.customer_id = os.customer_id
      where os.customer_id is not null
      and ss.customer_id is not null
      group by ss.customer_id")

```

#### 3. Write code to provide the count of distinct clients, total orders, total sales, and average purchase by purchase channel
```{r}

sqldf("select 
          purchase_channel
          ,count(distinct customer_id) as distinct_clients
          ,count(distinct order_number) as total_orders
          ,sum(total_amt) as total_sales
          ,avg(total_amt) as avg_purchase
      from line_sales 
      group by purchase_channel")



```

#### 4.Write code to provide a list of distinct clients that have an email address, signed up for email after 01/01/2010, and purchased tops after 3/10/2010
**Note: sqldf package in R does not like the date filters. This will run in a SQL editor**
```{r error=TRUE, eval=FALSE}


sqldf("select 
        c.customer_id
      from d_customer c
      join line_sales ls 
        on c.customer_id = ls.customer_id
      join product_table pt
        on ls.product_key on pt.product_key
      where c.email_address is not null 
      and c.signup_date > '2010-01-01'
      and ls.order_date > '2010-03-10'
      and pt.class = 'tops' ")

```
#### 5.Write code to provide a list of each email address and each email addresses total sales.
```{r}


sqldf("select 
        c.email_address, 
        sum(ls.total_amt) as total_sales
      from d_customer c 
      join line_sales ls
        on c.customer_id = ls.customer_id
      where c.email_address is not null
      group by c.email_address")
```
